Basic concepts and tools for the Toki Pona minimalist language: Wordnet synsets; analysis, synthesis and syntax highlighting of texts

### 2017-11-24 article/article.tex
andando com as estatisticas

% my view
% considerations about the official book and jan Pije
% pictorial forms
\section{Software for analysis and synthesis and for syntax highlighting}\label{hacks}
In this section,
we describe software, natural language, audiovisual and numeric results
on four directions:
1)
the analysis of the Toki Pona language through statistics
obtained by processing the dictionary;
2)
the synthesis of sentences, poems and short stories
in Toki Pona;
3)
syntax highlighting for Toki Pona,
including fine-tuning and theoretical
considerations;
4)
initial considerations about the WordNet of
Toki Pona words.
Corpus analysis was already found in~\cite{}.

@@

As described in Section~\cite{basics},
there is only 14 letters,
and phonemes respect a few rules.
There are 120 different words in the official vocabulary,
4 of them having synonyms.
A total of 124 tokens.
Not counting proper nouns (names)
and punctuation.
Table~\ref{tab:pos} shows the number of words
related to each POS tag specified in the dictionary.

@@

@@

>>>>

% my view
% considerations about the official book and jan Pije
% pictorial forms
\section{Software for analysis, synthesis, and syntax highlighting}\label{hacks}
In this section,
we describe software, natural language, audiovisual and numeric results
on four directions:
1)
the analysis of the Toki Pona language through statistics
obtained by processing the dictionary;
2)
the synthesis of sentences, poems and short stories
in Toki Pona;
3)
syntax highlighting for Toki Pona,
including fine-tuning and theoretical
considerations;
4)
initial considerations about the WordNet of
Toki Pona words.
Corpus analysis was already found in~\cite{}.

@@

As described in Section~\cite{basics},
there are only 14 letters,
and phonemes respect a few rules.
There are 120 different words in the official vocabulary,
4 of them having synonyms.
A total of 124 tokens,
not counting proper nouns (names)
and punctuation.
Table~\ref{tab:pos} shows the number of words
related to each POS tag specified in the dictionary.

@@

From the official 124 words, 
26 of them (20.97\%)
have only one syllable,
85 (68.55\%) have two syllables,
and 13 (10.48\%) have three syllables.
No official words has four or more syllables.
Of all the 235 syllables of the official words,
li, la, ka, na, and pa are the most often (with 5.53\%,
4.26\%, 3.83\%,  3.83\%, 3.83\%).
On the first syllable of each word,
a, o, pi, ka, and la are the most often
(with 6.45\%, 4.03\%, 4.03\%, 3.23\%, 3.23\%).
On the last syllable of each word,
li, lo, na, la, and ma are the most often
(with 8.06\%, 4.84\%, 4.84\%, 4.03\%, 4.03\%).
Middle-word syllables only occur 13 times,
and are all different, with the exception of `la',
which occurs twice.
For the frequencies of all syllables in every position,
run \tttt{makeStatistics.py} in~\cite{tokipona}.

@@

\input{vowels}
